---
title: "Data 607 Tidyverse Create"
author: "Kim Koon"
date: "`r Sys.Date()`"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

### Load packages

```{r load-packages, message=FALSE}
library(tidyverse)
library(sqldf)
options(scipen=999)
```

### Overview / Introduction

When creating outputs in SQL, I find myself using functions such as rank() over partition() and case when/then/end very frequently.  In this vignette, I will describe equivalents in tidyverse, using a Civil Engineering Salary survey sourced from the Civil Engineering subreddit.  

Imagine two scenarios where we would want to determine the highest reported salary by State, and determine whether a reported salary is lower or higher than the State average.  For scenario 2, since the dataset has over 2000 records, I will return only results from New York.

## Load/Transform Data

After importing the data, I filter on select columns of interest and rename the columns to be less verbose.  I also export this table into a CSV file for later use in SQL server when looking at SQL equivalents.

```{r load-data}
CE_Salary <- read.csv("C:\\Users\\Kim\\Documents\\GitHub\\Data607\\TidyVerse CREATE\\2023_CE_Salary_Survey.csv", header = TRUE, sep = ",")
CE_Salary <- CE_Salary %>% filter(Please.select.your.country. == 'United States of America') %>% select(What.state.do.you.work.in., Select.your.highest.level.of.education.attainment,Select.your.highest.level.of.professional.attainment..1, Which.best.describes.your.participation.in.the.Civil.Engineering.industry., Which.sub.discipline.do.you.work.in., Do.you.work.in.the.Public.or.Private.sector., What.is.your.current.annual.salary.) %>% rename(State = What.state.do.you.work.in., Education = Select.your.highest.level.of.education.attainment, Certification = Select.your.highest.level.of.professional.attainment..1, Partitipation = Which.best.describes.your.participation.in.the.Civil.Engineering.industry., Subdiscipline = Which.sub.discipline.do.you.work.in., Sector = Do.you.work.in.the.Public.or.Private.sector., Salary = What.is.your.current.annual.salary.)

write.table(CE_Salary, "C:\\Users\\Kim\\Documents\\GitHub\\Data607\\TidyVerse CREATE\\CE_Salary.csv", quote = FALSE, row.names = FALSE, sep = ";")
```

## SQL

To achieve the above aforementioned goals, I would use common table expressions.  I first create the CTE "stateRankedSalaries" to find the rank of each reported salary in comparison to other records for the state, sorted in descending order such that highest salaries are ranked first.  Then, I select from the CTE where the rank is equal to 1.  

I then create a CTE "avgStateSalaries" to find the average salary grouped by State.  I select from this CTE and include a new column "Compare to Average" where if the avgStateSalary is greater than the Salary of the record, it returns "Under", if the avgStateSalary is less than the Salary of the record, it returns "Above", and if they are equal, it returns "Average".  The state is filtered to "New York" in the where clause.

```{SQL, eval = FALSE}
-- preventing table conflicts
drop table if exists CE_Salary

-- creating the table
Create table CE_Salary (
State varchar(50),
Education varchar(50),
Certification varchar(50),
Participation varchar(50),
Subdiscipline varchar(max),
Sector varchar(50),
Salary float)

-- inserting the table from csv
bulk insert CE_Salary from 'C:\Users\Kim\Documents\GitHub\Data607\TidyVerse CREATE\CE_Salary.csv' with (firstrow = 2, fieldterminator = ';', rowterminator = '\n');


with stateRankedSalaries as (
  select *, rank() over (partition by State order by cast(Salary as int) desc) as salaryRank from CE_Salary 
)
Select * from stateRankedSalaries where salaryRank = 1
order by Salary desc;

with avgStateSalaries as (
	select *, avg(Salary) over (partition by State) as avgStateSalary from CE_Salary
	)
select *, case when avgStateSalary > Salary then 'Under' 
				when avgStateSalary < Salary then 'Above'
				when avgStateSalary = Salary then 'Average' end as 'Compare to Average'
				from avgStateSalaries
where State = 'New York'
```

To do the same in R, I first use the group_by() function to group by state, and then use mutate() with dense_rank() on the Salary in descending order using desc().  This combination of actions is the equivalent of rank() over (partition by State order by Salary desc) in SQL.  Then, I filter the dataset using filter() where the salaryRank is equal to 1, equivalent to the where clause in SQL.  I use arrange() to arrange the records in descending salary order, where arrange() is the equivalent to order by in SQL.

More interesting is the case_when() function in R, which allows multiple if/else statements like case when in SQL.  Using mutate with case_when(), I can perform the exact same process as in SQL and create different if/else scenarios that return specified outputs.  

```{r}
stateRankedSalaries <- CE_Salary %>% group_by(State) %>% mutate(salaryRank = dense_rank(desc(Salary))) %>% filter(salaryRank == 1) %>% arrange(desc(Salary))
print(stateRankedSalaries, n = 100)

avgStateSalaries <- CE_Salary %>% group_by(State) %>% mutate(avgStateSalary = mean(Salary)) %>% 
  mutate(CompareToAverage = case_when(avgStateSalary > Salary ~ "Under",
                   avgStateSalary < Salary ~ "Above",
                   avgStateSalary == Salary ~ "Average")) %>%
  filter(State == "New York")
print(avgStateSalaries, n = 150)
```

## Conclusion

The tidyverse packages, specifically dplyr, have various functions that can be used to perform equivalent SQL actions in R.  